---
title: "Final Code - R Era"
author: "Jeda Krisnell Dionisio"
date: "2026-05-28"
output: pdf_document
---

# Final Code
Below is the code used for the Rackauckas Era (2013-2017) during his time as DA. This Rmd analyzes the Black-White Becker hit rate based on if the gang enhancement was sustained. 

## Libraries + Data

```{r}
# ===== libraries =====
library(tidyverse)
library(ggplot2)

# ===== data =====
## ===== RACKAUCKAS ======
dat2013 = read.csv("../185CLEANED/2013-FINAL-CLEANED.csv")
dat2014 = read.csv("../185CLEANED/2014-FINAL-CLEANED.csv")
dat2015 = read.csv("../185CLEANED/2015-FINAL-CLEANED.csv")
dat2016 = read.csv("../185CLEANED/2016-FINAL-CLEANED.csv")
dat2017 = read.csv("../185CLEANED/2017-FINAL-CLEANED.csv")
```

# Black White Only
For the purposes of this analysis, I am only looking at Black and White. This code block makes it so White is the reference `Race`.

```{r}
# setting the reference groups
reference_data = function(data) {
  data |> 
    filter(Race %in% c("Black", "White")) |> 
    filter(!is.na(Disposition_enhancement)) |> 
    mutate(
      Race = relevel(factor(Race), ref = "White")
    )
}

dat2013 = reference_data(dat2013)
dat2014 = reference_data(dat2014)
dat2015 = reference_data(dat2015)
dat2016 = reference_data(dat2016)
dat2017 = reference_data(dat2017)
```

# Race Distribution
This code block gives the number of Black and White defendants. This is significant to account for the hit rate.

```{r}
table(dat2013$Race) 
table(dat2014$Race)
table(dat2015$Race)
table(dat2016$Race)
table(dat2017$Race)
```

# Get 'sustained'
I flag the `Disposition_enhancement` column since it describes if the enhancement was True or False. If the enhancement was passed, then it is flagged as 1 (`sustained`).

```{r}
create_sustained = function(data) {
  data |> 
    mutate(
      sustained = ifelse(
        Disposition_enhancement %in% c("True", "TRUE"),
        1,
        0
      )
    )
}


dat2013 = create_sustained(dat2013)
dat2014 = create_sustained(dat2014)
dat2015 = create_sustained(dat2015)
dat2016 = create_sustained(dat2016)
dat2017 = create_sustained(dat2017)
```

# Proportion Tables
Here is the raw Becker hit rate. 

```{r}
sustained_rate_table = function(data) {
  data |>
    filter(Race %in% c("Black", "White")) |>
    group_by(Race) |>
    summarise(
      n = n(),
      sustains = sum(sustained, na.rm = TRUE),
      misses = n() - sustains,
      sustained_rate = mean(sustained, na.rm = TRUE),
      .groups = "drop"
    )
}

bind_rows(
  sustained_rate_table(dat2013) |> mutate(year = 2013),
  sustained_rate_table(dat2014) |> mutate(year = 2014),
  sustained_rate_table(dat2015) |> mutate(year = 2015),
  sustained_rate_table(dat2016) |> mutate(year = 2016),
  sustained_rate_table(dat2017) |> mutate(year = 2017)
) |>
  select(year, everything())
```

# Logistic Regressions
Below are the logistic regression tests per year. One model predicts `sustained` based on `Race`, while the second model adds `has_prior`. Explanations of the models will be in the paper.

## 2013

### Model 1
```{r}
mod1 = glm(sustained ~ Race, 
           family = "binomial", 
           data = dat2013)
summary(mod1)
```

### Model 2
```{r}
mod2 = glm(sustained ~ Race + has_prior, 
           family = "binomial", 
           data = dat2013)
summary(mod2)
```

## 2014

### Model 3
```{r}
mod3 = glm(sustained ~ Race, 
           family = "binomial", 
           data = dat2014)
summary(mod3)
```

### Model 4
```{r}
mod4 = glm(sustained ~ Race + has_prior, 
           family = "binomial", 
           data = dat2014)
summary(mod4)
```

## 2015

### Model 5
```{r}
mod5 = glm(sustained ~ Race, 
           family = "binomial", 
           data = dat2015)
summary(mod5)
```

### Model 6
```{r}
mod6 = glm(sustained ~ Race + has_prior, 
           family = "binomial", 
           data = dat2015)
summary(mod6)
```

## 2016

### Model 7
```{r}
mod7 = glm(sustained ~ Race, 
           family = "binomial", 
           data = dat2016)
summary(mod7)
```

### Model 8
```{r}
mod8 = glm(sustained ~ Race + has_prior, 
           family = "binomial", 
           data = dat2016)
summary(mod8)
```

## 2017

### Model 9
```{r}
mod9 = glm(sustained ~ Race, 
           family = "binomial", 
           data = dat2017)
summary(mod9)
```

### Model 10
```{r}
mod10 = glm(sustained ~ Race + has_prior, 
           family = "binomial", 
           data = dat2017)
summary(mod10)
```