---
title: "Cleaning OCPD"
author: "Jeda Krisnell Dionisio"
date: "2026-05-28"
output: pdf_document
---

# Cleaning

Below is the code I used for the final project. I cleaned the data using this file to prevent the environment from getting too cluttered. For the sake of temporal analysis, though this is not the primary investigation, I split the years into the Rackauckas Era (2013-2017) and Spitzer Era (2018-2019, 2021, 2023-2024).

## Libraries + Files

```{r}
# ===== libraries =====
library(tidyverse)
library(readxl)
```

### RACKAUCKAS ERA

```{r}
# ===== 2013 =====
charges_filed2013 = read_excel("2013DATA/2013CHARGE.xlsx", sheet = "Data")
enhancements_charged2013 = read_excel("2013DATA/2013Enhancement.xlsx", sheet = "Data")
priors_alleged2013 = read_excel("2013DATA/2013Prior.xlsx", sheet = "Data")
sentences2013 = read_excel("2013DATA/2013Sentence.xlsx", sheet = "Data")
defendant_information2013 = read_excel("2013DATA/2013Def.xlsx", sheet = "Data")

# ===== 2014 =====
charges_filed2014 = read_excel("2014DATA/2014CHARGE.xlsx", sheet = "Data")
enhancements_charged2014 = read_excel("2014DATA/2014Enhancement.xlsx", sheet = "Data")
priors_alleged2014 = read_excel("2014DATA/2014Prior.xlsx", sheet = "Data")
sentences2014 = read_excel("2014DATA/2014Sentence.xlsx", sheet = "Data")
defendant_information2014 = read_excel("2014DATA/2014Def.xlsx", sheet = "Data")

# ===== 2015 =====
charges_filed2015 = read_excel("2015DATA/2015-Charges-Filed.xlsx", sheet = "Data")
enhancements_charged2015 = read_excel("2015DATA/2015-Enhancements-Charged.xlsx", sheet = "Data")
priors_alleged2015 = read_excel("2015DATA/2015-Priors-Alleged.xlsx", sheet = "Data")
sentences2015 = read_excel("2015DATA/2015-Sentences.xlsx", sheet = "Data")
defendant_information2015 = read_excel("2015DATA/2015-Defendant-Information.xlsx", sheet = "Data")

# ===== 2016 =====
charges_filed2016 = read_excel("2016DATA/2016-Charges-Filed.xlsx", sheet = "Data")
enhancements_charged2016 = read_excel("2016DATA/2016-Enhancements-Charged.xlsx", sheet = "Data")
priors_alleged2016 = read_excel("2016DATA/2016-Priors-Alleged.xlsx", sheet = "Data")
sentences2016 = read_excel("2016DATA/2016-Sentences.xlsx", sheet = "Data")
defendant_information2016 = read_excel("2016DATA/2016-Defendant-Information.xlsx", sheet = "Data")

# ===== 2017 =====
charges_filed2017 = read_excel("2017DATA/2017-Charges-Filed.xlsx", sheet = "Data")
enhancements_charged2017 = read_excel("2017DATA/2017-Enhancements-Charged.xlsx", sheet = "Data")
priors_alleged2017 = read_excel("2017DATA/2017-Priors-Alleged.xlsx", sheet = "Data")
sentences2017 = read_excel("2017DATA/2017-Sentences.xlsx", sheet = "Data")
defendant_information2017 = read_excel("2017DATA/2017-Defendant-Information.xlsx", sheet = "Data")

```

### SPITZER ERA

```{r}
# ===== 2018 =====
charges_filed2018 = read_excel("2018DATA/2018-Charges.xlsx", sheet = "Data")
enhancements_charged2018 = read_excel("2018DATA/2018-Enhancement.xlsx", sheet = "Data")
priors_alleged2018 = read_excel("2018DATA/2018-Prior.xlsx", sheet = "Data")
sentences2018 = read_excel("2018DATA/2018-Sentence.xlsx", sheet = "Data")
defendant_information2018 = read_excel("2018DATA/2018-Defendant.xlsx", sheet = "Data")

# ===== 2019 =====
charges_filed2019 = read_excel("2019DATA/2019-Charges-Filed.xlsx", sheet = "Data")
enhancements_charged2019 = read_excel("2019DATA/2019-Enhancements-Charged.xlsx", sheet = "Data")
priors_alleged2019 = read_excel("2019DATA/2019-Priors-Alleged.xlsx", sheet = "Data")
sentences2019 = read_excel("2019DATA/2019-Sentences.xlsx", sheet = "Data")
defendant_information2019 = read_excel("2019DATA/2019-Defendant-Information.xlsx", sheet = "Data")

# ===== 2021 =====
charges_filed2021 = read_excel("2021DATA/2021-Charges.xlsx", sheet = "Data")
enhancements_charged2021 = read_excel("2021DATA/2021-Enhancement.xlsx", sheet = "Data")
priors_alleged2021 = read_excel("2021DATA/2021-Prior.xlsx", sheet = "Data")
sentences2021 = read_excel("2021DATA/2021-Sentence.xlsx", sheet = "Data")
defendant_information2021 = read_excel("2021DATA/2021-Def.xlsx", sheet = "Data")

# ===== 2023 =====
charges_filed2023 = read_excel("2023DATA/2023Charges-Filed.xlsx", sheet = "Data")
enhancements_charged2023 = read_excel("2023DATA/2023Enhancements-Charged.xlsx", sheet = "Data")
priors_alleged2023 = read_excel("2023DATA/2023Prior.xlsx", sheet = "Data")
sentences2023 = read_excel("2023DATA/2023Sentences.xlsx", sheet = "Data")
defendant_information2023 = read_excel("2023DATA/2023Defendant-Information.xlsx", sheet = "Data")

# ===== 2024 =====
charges_filed2024 = read_excel("2024DATA/2024Charges.xlsx", sheet = "Data")
enhancements_charged2024 = read_excel("2024DATA/2024Enhancement.xlsx", sheet = "Data")
priors_alleged2024 = read_excel("2024DATA/2024Prior.xlsx", sheet = "Data")
sentences2024 = read_excel("2024DATA/2024Sentence.xlsx", sheet = "Data")
defendant_information2024 = read_excel("2024DATA/2024Def.xlsx", sheet = "Data")
```

# Cleaning

## RACKAUCKAS ERA

### 2013

```{r}
# ===== 2013 ======

# ===== filter gang enhancements =====
dat2013 = enhancements_charged2013 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2013 = dat2013 |> 
  left_join(defendant_information2013, by = "Defendant Number")

# ===== add primary charges ===== 
dat2013 = dat2013 |> 
  left_join(charges_filed2013,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2013_clean = sentences2013 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2013 = dat2013 |> 
  left_join(
    sentences2013_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2013 = priors_alleged2013 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2013 = dat2013 |> 
  left_join(priors_summary2013,
            by = c("Case ID Number", "Defendant Number"))

dat2013 = dat2013 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2013 = dat2013 |> 
  filter(Gender != "Unknown")

dim(dat2013)

# ===== save dataframe to use in Final ===== 
write.csv(dat2013, "../185CLEANED/2013-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2014

```{r}
# ===== 2014 ======

# ===== filter gang enhancements =====
dat2014 = enhancements_charged2014 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2014 = dat2014 |> 
  left_join(defendant_information2014, by = "Defendant Number")

# ===== add primary charges ===== 
dat2014 = dat2014 |> 
  left_join(charges_filed2014,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2014_clean = sentences2014 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2014 = dat2014 |> 
  left_join(
    sentences2014_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2014 = priors_alleged2014 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2014 = dat2014 |> 
  left_join(priors_summary2014,
            by = c("Case ID Number", "Defendant Number"))

dat2014 = dat2014 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2014 = dat2014 |> 
  filter(Gender != "Unknown")

dim(dat2014)

# ===== save dataframe to use in Final ===== 
write.csv(dat2014, "../185CLEANED/2014-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2015

```{r}
# ===== 2015 ======

# ===== filter gang enhancements =====
dat2015 = enhancements_charged2015 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2015 = dat2015 |> 
  left_join(defendant_information2015, by = "Defendant Number")

# ===== add primary charges ===== 
dat2015 = dat2015 |> 
  left_join(charges_filed2015,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2015_clean = sentences2015 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2015 = dat2015 |> 
  left_join(
    sentences2015_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2015 = priors_alleged2015 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2015 = dat2015 |> 
  left_join(priors_summary2015,
            by = c("Case ID Number", "Defendant Number"))

dat2015 = dat2015 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2015 = dat2015 |> 
  filter(Gender != "Unknown")

dim(dat2015)

# ===== save dataframe to use in Final ===== 
write.csv(dat2015, "../185CLEANED/2015-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2016

```{r}
# ===== 2016 ======

# ===== filter gang enhancements =====
dat2016 = enhancements_charged2016 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2016 = dat2016 |> 
  left_join(defendant_information2016, by = "Defendant Number")

# ===== add primary charges ===== 
dat2016 = dat2016 |> 
  left_join(charges_filed2016,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2016_clean = sentences2016 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2016 = dat2016 |> 
  left_join(
    sentences2016_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2016 = priors_alleged2016 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2016 = dat2016 |> 
  left_join(priors_summary2016,
            by = c("Case ID Number", "Defendant Number"))

dat2016 = dat2016 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2016 = dat2016 |> 
  filter(Gender != "Unknown")

dim(dat2016)

# ===== save dataframe to use in Final ===== 
write.csv(dat2016, "../185CLEANED/2016-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2017

```{r}
# ===== 2017 ======

# ===== filter gang enhancements =====
dat2017 = enhancements_charged2017 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2017 = dat2017 |> 
  left_join(defendant_information2017, by = "Defendant Number")

# ===== add primary charges ===== 
dat2017 = dat2017 |> 
  left_join(charges_filed2017,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2017_clean = sentences2017 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2017 = dat2017 |> 
  left_join(
    sentences2017_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2017 = priors_alleged2017 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2017 = dat2017 |> 
  left_join(priors_summary2017,
            by = c("Case ID Number", "Defendant Number"))

dat2017 = dat2017 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2017 = dat2017 |> 
  filter(Gender != "Unknown")

dim(dat2017)

# ===== save dataframe to use in Final ===== 
write.csv(dat2017, "../185CLEANED/2017-FINAL-CLEANED.csv", row.names = FALSE)
```

## SPITZER ERA

### 2018

```{r}
# ===== 2018 ======

# ===== filter gang enhancements =====
dat2018 = enhancements_charged2018 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2018 = dat2018 |> 
  left_join(defendant_information2018, by = "Defendant Number")

# ===== add primary charges ===== 
dat2018 = dat2018 |> 
  left_join(charges_filed2018,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2018_clean = sentences2018 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count) 

# ====== join cleaned sentencing =====
dat2018 = dat2018 |> 
  left_join(
    sentences2018_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2018 = priors_alleged2018 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2018 = dat2018 |> 
  left_join(priors_summary2018,
            by = c("Case ID Number", "Defendant Number"))

dat2018 = dat2018 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2018 = dat2018 |> 
  filter(Gender != "Unknown")

dim(dat2018)

# ===== save dataframe to use in Final ===== 
write.csv(dat2018, "../185CLEANED/2018-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2019

```{r}
# ===== 2019 ======

# ===== filter gang enhancements =====
dat2019 = enhancements_charged2019 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2019 = dat2019 |> 
  left_join(defendant_information2019, by = "Defendant Number")

# ===== add primary charges ===== 
dat2019 = dat2019 |> 
  left_join(charges_filed2019,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2019_clean = sentences2019 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2019 = dat2019 |> 
  left_join(
    sentences2019_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2019 = priors_alleged2019 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2019 = dat2019 |> 
  left_join(priors_summary2019,
            by = c("Case ID Number", "Defendant Number"))

dat2019 = dat2019 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2019 = dat2019 |> 
  filter(Gender != "Unknown")

dim(dat2019)

# ===== save dataframe to use in Final ===== 
write.csv(dat2019, "../185CLEANED/2019-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2021

```{r}
# ===== 2021 ======

# ===== filter gang enhancements =====
dat2021 = enhancements_charged2021 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2021 = dat2021 |> 
  left_join(defendant_information2021, by = "Defendant Number")

# ===== add primary charges ===== 
dat2021 = dat2021 |> 
  left_join(charges_filed2021,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2021_clean = sentences2021 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2021 = dat2021 |> 
  left_join(
    sentences2021_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2021 = priors_alleged2021 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2021 = dat2021 |> 
  left_join(priors_summary2021,
            by = c("Case ID Number", "Defendant Number"))

dat2021 = dat2021 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2021 = dat2021 |> 
  filter(Gender != "Unknown")

dim(dat2021)

# ===== save dataframe to use in Final ===== 
write.csv(dat2021, "../185CLEANED/2021-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2023

```{r}
# ===== 2023 ======

# ===== filter gang enhancements =====
dat2023 = enhancements_charged2023 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2023 = dat2023 |> 
  left_join(defendant_information2023, by = "Defendant Number")

# ===== add primary charges ===== 
dat2023 = dat2023 |> 
  left_join(charges_filed2023,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2023_clean = sentences2023 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count) 

# ====== join cleaned sentencing =====
dat2023 = dat2023 |> 
  left_join(
    sentences2023_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2023 = priors_alleged2023 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2023 = dat2023 |> 
  left_join(priors_summary2023,
            by = c("Case ID Number", "Defendant Number"))

dat2023 = dat2023 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2023 = dat2023 |> 
  filter(Gender != "Unknown")

dim(dat2023)

# ===== save dataframe to use in Final ===== 
write.csv(dat2023, "../185CLEANED/2023-FINAL-CLEANED.csv", row.names = FALSE)
```

### 2024

```{r}
# ===== 2024 ======

# ===== filter gang enhancements =====
dat2024 = enhancements_charged2024 |> 
  filter(str_detect(`Offense Code`, "186.22"))

# ===== add defendant information ===== 
dat2024 = dat2024 |> 
  left_join(defendant_information2024, by = "Defendant Number")

# ===== add primary charges ===== 
dat2024 = dat2024 |> 
  left_join(charges_filed2024,
            by = c("Case ID Number", "Defendant Number", "Count"),
            suffix = c("_enhancement", "_primary"))

# ===== clean sentences =====
sentences2024_clean = sentences2024 |> 
  distinct(`Case ID Number`, `Defendant Number`, Count)

# ====== join cleaned sentencing =====
dat2024 = dat2024 |> 
  left_join(
    sentences2024_clean,
    by = c("Case ID Number", "Defendant Number", "Count")
  )

# ===== create priors_alleged summary table ===== 
priors_summary2024 = priors_alleged2024 |> 
  group_by(`Case ID Number`, `Defendant Number`) |> 
  summarise(
    has_prior = 1,
    .groups = "drop"
  )

# ===== merge priors into main dat ===== 
dat2024 = dat2024 |> 
  left_join(priors_summary2024,
            by = c("Case ID Number", "Defendant Number"))

dat2024 = dat2024 |> 
  mutate(
    has_prior = ifelse(is.na(has_prior), 0, has_prior)  
  )

# ===== final cleaning ===== 
dat2024 = dat2024 |> 
  filter(Gender != "Unknown")

dim(dat2024)

# ===== save dataframe to use in Final ===== 
write.csv(dat2024, "../185CLEANED/2024-FINAL-CLEANED.csv", row.names = FALSE)
```
